Defocus Corrected Large Area Cryo-EM (DeCo-LACE) for Label-Free Detection of Molecules across Entire Cell Sections

### 2022-06-14 manuscript.tex
Place movies into html version [ci skip]
This build is based on
https://github.com/jojoelfe/deco_lace_template_matching_manuscript/commit/009ecfe7c505910c424b8f1b31b59f44373441da.

This commit was created by the following CI build and job:
https://github.com/jojoelfe/deco_lace_template_matching_manuscript/commit/009ecfe7c505910c424b8f1b31b59f44373441da/checks
https://github.com/jojoelfe/deco_lace_template_matching_manuscript/actions/runs/2491907565

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Defocus Corrected Large Area Cryo-EM (DeCo-LACE) for Label-Free Detection of Molecules across Entire Cell Sections},
  pdfauthor={Johannes ElferichGiulia SchiroliDavid ScaddenNikolaus Grigorieff},
  pdflang={en-US},
  pdfkeywords={cryo-EM, visual protemics, ribosome},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{american}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\usepackage[margin=2cm]{geometry}
\usepackage{setspace}
\doublespacing
\usepackage{lineno}
\linenumbers
\usepackage{cleveref}

@@

This manuscript
(\href{https://jojoelfe.github.io/deco_lace_template_matching_manuscript/v/a321def4b341f6e6f2c8d66936f53cf5e9398eae/}{permalink})
was automatically generated
from \href{https://github.com/jojoelfe/deco_lace_template_matching_manuscript/tree/a321def4b341f6e6f2c8d66936f53cf5e9398eae}{jojoelfe/deco\_lace\_template\_matching\_manuscript@a321def}
on June 6, 2022.

@@

\begin{itemize}
\item
  \textbf{Johannes Elferich}
  \includegraphics[width=0.16667in,height=0.16667in]{images/orcid.pdf}
  \href{https://orcid.org/0000-0002-9911-706X}{0000-0002-9911-706X}
  · \includegraphics[width=0.16667in,height=0.16667in]{images/github.pdf}
  \href{https://github.com/jojoelfe}{jojoelfe}
  RNA Therapeutics Institute, University of Massachusetts Chan Medical School, Worcester, MA, USA; Howard Hughes Medical Institute, Ashburn, VA, USA
\item
  \textbf{Giulia Schiroli}
  \includegraphics[width=0.16667in,height=0.16667in]{images/orcid.pdf}
  \href{https://orcid.org/0000-0002-9812-8728}{0000-0002-9812-8728}
  Center for Regenerative Medicine, Massachusetts General Hospital, Boston, MA 02114, USA
\item
  \textbf{David Scadden}
  Center for Regenerative Medicine, Massachusetts General Hospital, Boston, MA 02114, USA; Harvard Stem Cell Institute, Cambridge, MA 02138, USA; Department of Stem Cell and Regenerative Biology, Harvard University, Cambridge, MA 02138, USA
\item
  \textbf{Nikolaus Grigorieff}
  \includegraphics[width=0.16667in,height=0.16667in]{images/orcid.pdf}
  \href{https://orcid.org/0000-0002-1506-909X}{0000-0002-1506-909X}
  · \includegraphics[width=0.16667in,height=0.16667in]{images/github.pdf}
  \href{https://github.com/nikogrigorieff}{nikogrigorieff}
  RNA Therapeutics Institute, University of Massachusetts Chan Medical School, Worcester, MA, USA; Howard Hughes Medical Institute, Ashburn, VA, USA
\end{itemize}

@@

Here we apply 2DTM of the ribosome large subunit (LSU) to conditionally immortalized \emph{mus musculus} (mouse) cell line that gives rise to functional mature neutrophils {[}\protect\hyperlink{ref-1B9Vt9eYu}{10}{]} after thinning using cryo-FIB milling. We chose these
cells because genetic defects in the ribosome machinery often leads to
hematopoietic disease {[}\protect\hyperlink{ref-gRoY21jY}{11}{]} and direct quantification of
ribosome location, number and conformational states in hematopoietic cells could
lead to new insight into hematopoietic disease {[}\protect\hyperlink{ref-KAJ7221k}{12}{]}.
To increase the amount of collected data and to provide unbiased sampling of the
whole lamella, we devised a new data-acquisition scheme, ``Defocus-Corrected
Large Area Cryo-Electron microscopy'' (DeCo-LACE). 2DTM allows us to test whether
aberrations caused by large beam-image shifts and highly condensed beams
deteriorate the high-resolution signal. We find that these aberrations do not
impede LSU detection by 2DTM. The resulting data provide a description of
ribosome distribution in an entire lamella, which represent 1-2\% of the cellular
volume. We find a highly heterogeneous density of ribosomes within the cell.
Analysis of the throughput in this method suggests that for the foreseeable
future computation will be the bottleneck for visual proteomics.

@@

We used the tile positions to calculate the positions of the detected
LSUs in the lamellae (Fig. {[}\ref{fig:matching_euc}{]}A, Fig. {[}\ref{fig:matching_fff}{]}A). Overlaying these positions of the lamellae montages reveals LSU
distribution throughout the FIB-milled slices of individual cells. Consistent with prior observations imaging selected views in yeast {[}\protect\hyperlink{ref-T1Qr0xyP}{13}{]}, organelles
like the nucleus and mitochondria only showed sporadic targets detected with low
SNRs, consistent with the estimated false-positive rate of one per tile. For
each detected target we also calculated the Z positions from the individual
estimated defocus and defocus offset for each tile. When viewed from the side,
the ribosome positions therefore show the slight tilts of the lamellae relative
to the microscope frame of reference (Fig. {[}\ref{fig:matching_euc}{]}B, Fig.
{[}\ref{fig:matching_fff}{]}B). Furthermore, the side views indicated that lamellae
were thinner at the leading edge. Indeed, when plotting the transmitted beam
intensities in individual tiles as a function of beam image-shift, we observed
substantially higher intensities at the leading edge (Fig. {[}\ref{fig:lamella_motion_thickness}{]}B), which in energy-filtered
TEM indicates a thinner sample {[}\protect\hyperlink{ref-17IP4Lhz1}{18}{]}. Even though we prepared the lamellae
with the ``overtilt'' approach {[}\protect\hyperlink{ref-hr3dyUeI}{19}{]}, this means that LSU densities across
the lamellae can be skewed by a change in thickness, and better sample
preparation methods are needed to generate more even samples.

@@

Since the SNR values of 2DTM are highly sensitive to image quality, we reasoned
we could use them to verify that DeCo-LACE does not introduce a systematic loss
of image quality. We considered non-parallel illumination introduced by the
unusually condensed beam and uncharacterized aberrations near the beam
periphery. When plotting the SNR values of detected targets in all eight
lamellae as a function of their location in the tiles, we found uniformly high
SNR values throughout the illuminated areas for both eucentric and fringe-free
focus illumination, demonstrating that both illumination schemed are suitable
for DeCo-LACE (Fig. {[}\ref{fig:matching_stat}{]}D).

We also wondered whether large image shifts would lead aberration due to
astigmatism or beam tilt {[}\protect\hyperlink{ref-APSL9LmU}{16}{]}. We reasoned that if
that was the case the number of detected targets should be highest in the center of the
lamella where the applied beam image-shift is 0. Instead, we observed that in
both eucentric and fringe-free focus conditions more targets were detected at the
``back'' edge of the lamella (Fig. {[}\ref{fig:matching_stat}{]}E{]}). This may be due to the center of the cell being predominantly
occupied by the nucleus, despite its segmentation in neutrophil-like cells. The
increase in matches at the ``back'' of the lamellae compared to the ``front'' can also
be explained by the thickness gradient of the lamellae (Fig. {[}\ref{fig:lamella_motion_thickness}{]}B,
Fig. {[}\ref{fig:matching_euc}{]}B, Fig. {[}\ref{fig:matching_fff}{]}B). In addition, aberrations would be expected to cause average 2DTM SNRs to be higher when beam-image shift
values are small. Instead, we found that SNRs where on average the highest at the
``front'' edge of the lamellae, presumably due to the thinner sample. We therefore
conclude that factors other that beam image-shift or beam
condensation aberrations are limiting 2DTM SNRS, predominantly the thickness of
the lamellae.

@@

In this study we developed an approach to image entire cellular cross-section using cryo-EM at high enough resolution to allow for 2DTM detection of the LSU. The two main advantages compared to previous approaches are a high throughput of imaging and the biological context for detected molecules. The requirement to increase throughput in cryo-EM data collection of cellular samples has been recognized in the recent literature. Most approaches described so far are tailored towards tomography. Peck et al. {[}\protect\hyperlink{ref-1GPPSX111}{21}{]} and Yang et al. {[}\protect\hyperlink{ref-ujizcpVn}{22}{]} developed approaches to increase the FOV of tomogram data-collection by using a montaging technique. Peck et al.~used a similar ``condensed-beam'' approach as described here. However, the montages are substantially smaller in scope, covering carbon film holes of 2 um diameter. Bouvette et al. {[}\protect\hyperlink{ref-1Gwx5ScUa}{23}{]} and Eisenstein et al. {[}\protect\hyperlink{ref-xvouNMBy}{24}{]} are using beam image-shift to collect tilt-series in multiple locations in parallel to increase throughput. However, none of these approaches provide the full coverage of a cellular cross-section that can be achieved using DeCo-Lace.

Since we observed substantial variation in LSU density within and between lamellae, visual proteomics studies that use cryo-EM to establish changes in molecular organization within cells will require orders of magnitude more data than used in this study. One milestone would be to image enough data to represent one cellular volume, which for a small eukaryotic cells requires imaging approximately 100 lamella. While data collection throughput on the TEM is fundamentally limited by the exposure time, this amount of data could be collected within 12 hours by improving the data acquisition scheme to perform all necessary calculations in parallel with actual exposure of the camera. Sample preparation using a FIB/SEM is also currently a bottleneck, but preparation of large lamellae with multiple cellular cross-sections using methods like WAFFLE {[}\protect\hyperlink{ref-17H2LWiIu}{25}{]} might allow sufficient throughput. As stated in the results, at least for 2DTM computation will remain challenging and approximately 17,000 GPU hours would be required for a 100 lamellae dataset.

As described in {[}\protect\hyperlink{ref-Ynb3IP6I}{7}{]} the 2DTM SNR threshold for detecting a target is chosen to result in one false positive detection per image searched. We would therefore expect to find one false positive detection per tile. We reasoned that the large nuclear area imaged by DeCo-Lace could be used to test whether this assumption is true. In the 670 tiles containing exclusively nucleus (as manually annotated from the overview image) we detected 247 targets, making the false-positive rate more than twofold lower than expected. Since earlier work shows that 2DTM with the LSU can produce matches to nuclear ribosome biogenesis intermediates {[}\protect\hyperlink{ref-T1Qr0xyP}{13}{]}, this could even be an overestimate of the false-positive rate. This suggests that the detection threshold could be even lower, which is an area of ongoing research.

We found that even though we used beam image-shift extensively (up to 7 um), we did not see substantially reduced 2DTM SNR values in tiles acquired at high beam image-shift compared to tiles acquired with low or no beam image-shift. This is in contrast to reports in single-particle analysis (SPA) {[}\protect\hyperlink{ref-44fG6kcy}{26}{]} where the induced beam tilt substantially reduced the resolution if it was not corrected during processing. It is possible that 2DTM is less sensitive to beam-tilt aberrations, since the template is free of any aberration and only the image is distorted, while in SPA the beam tilt will affect both the images and the reconstructed template.

@@

We observed granules containing a vesicle of putative cytosolic origin. We speculate that upon degranulation, the process in which granules fuse with the plasma membrane, these vesicles would be released into the extracelullar space. The main types of extracellular vesicles of this size are exosomes, up to 100 nm large vesicles derived from fusion of multivesicular bodies with the plasma membrane, and microvesicles, which are derived from direct budding of the plasma membrane {[}\protect\hyperlink{ref-kncnYGLD}{27}{]}. We suggest that granulocytes could release a third type of extracellular vesicle, granule-derived vesicles (GDV), into the extracellular space. 2DTM showed that a subset of GDVs can contain ribosomes (Fig. {[}\ref{fig:matching_euc}{]}F, Fig. {[}\ref{fig:matching_fff}{]}D,E). This could indicate that these vesicles are transporting translation-capable mRNAs, as has been described for exosomes {[}\protect\hyperlink{ref-agVdpX76}{28}{]}. Further studies will be necessary to confirm the existence of GDVs in granulocytes isolated from mammals and to understand their functional significance.

@@

ER-HoxB8 cells were maintained in RPMI medium supplemented with 10\% FBS,
penicillin/streptomycin, SCF, and estrogen {[}\protect\hyperlink{ref-1B9Vt9eYu}{10}{]} at
37 °C and 5\% CO2. 120 h prior to grid preparation, cells were washed twice in PBS
and cultured in the same medium except without estrogen. Differentiation was
verified by staining with Hoechst-dye and inspection of nuclear morphology.
Cells were then counted and diluted to \(1\cdot10^6\) cells/ml. Grids (either 200
mesh copper grids, with a sillicone-oxide and 2 µm holes with a 2 µm spacing or
200 mesh gold grids with a thin gold film and 2 µm holes in 2 µm spacing) were
glow-discharged from both sides using a 15 mA for 45 s. 3.5 µl of cell suspension
was added to grids on the thin-film side and grids were blotted from the back
side using a GP2 cryoplunger (Leica) for 8 s and rapidly plunged into liquid
ethane at -185 °C.

@@

Grids were loaded into an Aquilos 2 FIB/SEM (Thermo Fisher) instrument with a
stage cooled to -190 °C. Grids were sputter-coated with platinum for 15 s at 45 mA
and then coated with a layer of platinum-precursor by opening the GIS-valve for
45 s. An overview of the grid was created by montaging SEM images and isolated
cells at the center of gridsquares were selected for FIB-milling. Lamellae were
generated automatically using the AutoTEM software (Thermo Fisher), with the
following parameters:

@@

After data collection, a 20 s exposure at 2250x magnification of the lamella at
200 μm defocus was taken for visualization purposes. A Python script implementing
this procedure is available at {[}Link to repo{]}.

@@

\begin{figure}
\hypertarget{fig:deco_lace_workflow}{%
\centering
\includegraphics{figures/deco_lace_workflow.png}
\caption{Workflow of DeCo-Lace processing}\label{fig:deco_lace_workflow}
}
\end{figure}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Defocus Corrected Large Area Cryo-EM (DeCo-LACE) for Label-Free Detection of Molecules across Entire Cell Sections},
  pdfauthor={Johannes ElferichGiulia SchiroliDavid ScaddenNikolaus Grigorieff},
  pdflang={en-US},
  pdfkeywords={cryo-EM, visual protemics, ribosome},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{american}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\usepackage[margin=2cm]{geometry}
\usepackage{setspace}
\doublespacing
\usepackage{lineno}
\linenumbers
\usepackage{cleveref}
\usepackage{fontspec}
\setmainfont{texgyrepagella}[
  Extension = .otf,
  UprightFont = *-regular,
  BoldFont = *-bold,
  ItalicFont = *-italic,
  BoldItalicFont = *-bolditalic,
]
\usepackage{float}
\floatplacement{figure}{H}
\usepackage{caption}
\captionsetup[figure]{font={stretch=1.2}}

@@

@@

\begin{itemize}
\item
  \textbf{Johannes Elferich}
  \includegraphics[width=0.16667in,height=0.16667in]{images/envelope.pdf}\includegraphics[width=0.16667in,height=0.16667in]{images/orcid.pdf}
  \href{https://orcid.org/0000-0002-9911-706X}{0000-0002-9911-706X}

  RNA Therapeutics Institute, University of Massachusetts Chan Medical School, Worcester, MA, USA; Howard Hughes Medical Institute, USA
\item
  \textbf{Giulia Schiroli}
  \includegraphics[width=0.16667in,height=0.16667in]{images/orcid.pdf}
  \href{https://orcid.org/0000-0002-9812-8728}{0000-0002-9812-8728}

  Center for Regenerative Medicine, Massachusetts General Hospital, Boston, MA, USA; Harvard Stem Cell Institute, Cambridge, MA, USA; Department of Stem Cell and Regenerative Biology, Harvard University, Cambridge, MA, USA
\item
  \textbf{David Scadden}
  \includegraphics[width=0.16667in,height=0.16667in]{images/orcid.pdf}
  \href{https://orcid.org/0000-0001-9821-7133}{0000-0001-9821-7133}

  Center for Regenerative Medicine, Massachusetts General Hospital, Boston, MA, USA; Harvard Stem Cell Institute, Cambridge, MA, USA; Department of Stem Cell and Regenerative Biology, Harvard University, Cambridge, MA, USA
\item
  \textbf{Nikolaus Grigorieff}
  \includegraphics[width=0.16667in,height=0.16667in]{images/envelope.pdf}\includegraphics[width=0.16667in,height=0.16667in]{images/orcid.pdf}
  \href{https://orcid.org/0000-0002-1506-909X}{0000-0002-1506-909X}

  RNA Therapeutics Institute, University of Massachusetts Chan Medical School, Worcester, MA, USA; Howard Hughes Medical Institute, USA
\end{itemize}

\leavevmode\vadjust pre{\hypertarget{correspondence}{}}%
\includegraphics[width=0.16667in,height=0.16667in]{images/envelope.pdf} --- correspondence preferred via \href{https://github.com/jojoelfe/deco_lace_template_matching_manuscript/issues}{GitHub Issues}.
Otherwise, address correspondence to \href{mailto:johannes.elferich@umassmed.edu}{\nolinkurl{johannes.elferich@umassmed.edu}} and \href{mailto:niko@grigorieff.org}{\nolinkurl{niko@grigorieff.org}}.

@@

Here we apply 2DTM of the ribosome large subunit (LSU) to a conditionally immortalized \emph{mus musculus} (mouse) cell line that gives rise to functional mature neutrophils {[}\protect\hyperlink{ref-1B9Vt9eYu}{10}{]}. We chose these
cells because genetic defects in the ribosome machinery often leads to
hematopoietic disease {[}\protect\hyperlink{ref-gRoY21jY}{11}{]} and direct quantification of
ribosome location, number and conformational states in hematopoietic cells could
lead to new insight into hematopoietic disease {[}\protect\hyperlink{ref-KAJ7221k}{12}{]}.
To increase the amount of collected data and to provide unbiased sampling of the
whole lamella, we devised a new data-acquisition scheme, ``Defocus-Corrected
Large Area Cryo-Electron microscopy'' (DeCo-LACE). 2DTM allows us to test whether
aberrations caused by large beam-image shifts and highly condensed beams
deteriorate the high-resolution signal. We find that these aberrations do not
impede LSU detection by 2DTM. The resulting data provide a description of
ribosome distribution in an entire lamella, which represent 1-2\% of the cellular
volume. We find a highly heterogeneous density of ribosomes within the cell.
Analysis of the throughput in this method suggests that for the foreseeable
future computation will be the bottleneck for visual proteomics.

@@

We used the tile positions to calculate the positions of the detected
LSUs in the lamellae (Fig. {[}\ref{fig:matching_euc}{]}A, Fig. {[}\ref{fig:matching_fff}{]}A, Movie 1, Movie 2). Overlaying these positions of the lamellae montages reveals LSU
distribution throughout the FIB-milled slices of individual cells. Consistent with prior observations imaging selected views in yeast {[}\protect\hyperlink{ref-T1Qr0xyP}{13}{]}, organelles
like the nucleus and mitochondria only showed sporadic targets detected with low
SNRs, consistent with the estimated false-positive rate of one per tile. For
each detected target we also calculated the Z positions from the individual
estimated defocus and defocus offset for each tile. When viewed from the side,
the ribosome positions therefore show the slight tilts of the lamellae relative
to the microscope frame of reference (Fig. {[}\ref{fig:matching_euc}{]}B, Fig.
{[}\ref{fig:matching_fff}{]}B, Movie 1, Movie 2). Furthermore, the side views indicated that lamellae
were thinner at the leading edge. Indeed, when plotting the transmitted beam
intensities in individual tiles as a function of beam image-shift, we observed
substantially higher intensities at the leading edge (Fig. {[}\ref{fig:lamella_motion_thickness}{]}B), which in energy-filtered
TEM indicates a thinner sample {[}\protect\hyperlink{ref-17IP4Lhz1}{18}{]}. Even though we prepared the lamellae
with the ``overtilt'' approach {[}\protect\hyperlink{ref-hr3dyUeI}{19}{]}, this means that LSU densities across
the lamellae can be skewed by a change in thickness, and better sample
preparation methods are needed to generate more even samples.

As described in {[}\protect\hyperlink{ref-Ynb3IP6I}{7}{]} the 2DTM SNR threshold for detecting a target is chosen to result in one false positive detection per image searched. We would therefore expect to find one false positive detection per tile. We reasoned that the large nuclear area imaged by DeCo-LACE could be used to test whether this assumption is true. In the 670 tiles containing exclusively nucleus (as manually annotated from the overview image) we detected 247 targets, making the false-positive rate more than twofold lower than expected. Since earlier work shows that 2DTM with the LSU can produce matches to nuclear ribosome biogenesis intermediates {[}\protect\hyperlink{ref-T1Qr0xyP}{13}{]}, this could even be an overestimate of the false-positive rate. This suggests that the detection threshold could be even lower, which is an area of ongoing research.

@@

Since the SNR values of 2DTM are highly sensitive to image quality, we reasoned
we could use them to verify that DeCo-LACE does not introduce a systematic loss
of image quality. We considered non-parallel illumination introduced by the
unusually condensed beam and uncharacterized aberrations near the beam
periphery. When plotting the SNR values of detected targets in all eight
lamellae as a function of their location in the tiles, we found uniformly high
SNR values throughout the illuminated areas for both eucentric and fringe-free
focus illumination, demonstrating that both illumination schemes are suitable
for DeCo-LACE (Fig. {[}\ref{fig:matching_stat}{]}D).

We also wondered whether large image shifts would lead to aberrations due to
astigmatism or beam tilt {[}\protect\hyperlink{ref-APSL9LmU}{16}{]}. We reasoned that if
that was the case the number of detected targets should be highest in the center of the
lamella where the applied beam image-shift is 0. Instead, we observed that in
both eucentric and fringe-free focus conditions more targets were detected at the
``back'' edge of the lamella (Fig. {[}\ref{fig:matching_stat}{]}E{]}). This may be due to the center of the cell being predominantly
occupied by the nucleus, despite its segmentation in neutrophil-like cells. The
increase in matches at the ``back'' of the lamellae compared to the ``front'' can also
be explained by the thickness gradient of the lamellae (Fig. {[}\ref{fig:lamella_motion_thickness}{]}B,
Fig. {[}\ref{fig:matching_euc}{]}B, Fig. {[}\ref{fig:matching_fff}{]}B). In addition, aberrations would be expected to cause average 2DTM SNRs to be higher when beam-image shift
values are small. Instead, we found that SNRs where on average the highest at the
``front'' edge of the lamellae, presumably due to the thinner sample. We therefore
conclude that factors other that beam image-shift or beam
condensation aberrations are limiting 2DTM SNRS, predominantly the thickness of
the lamellae.

@@

In this study we developed an approach to image entire cellular cross-section using cryo-EM at high enough resolution to allow for 2DTM detection of the LSU. The two main advantages compared to previous approaches are high throughput of imaging and biological context for detected molecules. The requirement to increase throughput in cryo-EM data collection of cellular samples has been recognized in the recent literature. Most approaches described so far are tailored towards tomography. Peck et al. {[}\protect\hyperlink{ref-1GPPSX111}{21}{]} and Yang et al. {[}\protect\hyperlink{ref-ujizcpVn}{22}{]} developed approaches to increase the FOV of tomogram data-collection by using a montaging technique. Peck et al.~used a similar ``condensed-beam'' approach as described here. However, the montages are substantially smaller in scope, covering carbon film holes of 2 µm diameter. Bouvette et al. {[}\protect\hyperlink{ref-1Gwx5ScUa}{23}{]} and Eisenstein et al. {[}\protect\hyperlink{ref-xvouNMBy}{24}{]} are using beam image-shift to collect tilt-series in multiple locations in parallel to increase throughput. However, none of these approaches provide the full coverage of a cellular cross-section that can be achieved using DeCo-LACE.

We observed granules containing a vesicle of putative cytosolic origin. We speculate that upon degranulation, the process in which granules fuse with the plasma membrane, these vesicles would be released into the extracelullar space. The main types of extracellular vesicles of this size are exosomes, up to 100 nm large vesicles derived from fusion of multivesicular bodies with the plasma membrane, and microvesicles, which are derived from direct budding of the plasma membrane {[}\protect\hyperlink{ref-kncnYGLD}{25}{]}. We suggest that granulocytes could release a third type of extracellular vesicle, granule-derived vesicles (GDV), into the extracellular space. 2DTM showed that a subset of GDVs can contain ribosomes (Fig. {[}\ref{fig:matching_euc}{]}F, Fig. {[}\ref{fig:matching_fff}{]}D,E). This could indicate that these vesicles are transporting translation-capable mRNAs, as has been described for exosomes {[}\protect\hyperlink{ref-agVdpX76}{26}{]}. Further studies will be necessary to confirm the existence of GDVs in granulocytes isolated from mammals and to understand their functional significance.

@@

We found that even though we used beam image-shift extensively (up to 7 um), we did not see substantially reduced 2DTM SNR values in tiles acquired at high beam image-shift compared to tiles acquired with low or no beam image-shift. This is in contrast to reports in single-particle analysis (SPA) {[}\protect\hyperlink{ref-44fG6kcy}{27}{]} where the induced beam tilt substantially reduced the resolution if it was not corrected during processing. It is possible that 2DTM is less sensitive to beam-tilt aberrations, since the template is free of any aberration and only the image is distorted, while in SPA the beam tilt will affect both the images and the reconstructed template.

Since we observed substantial variation in LSU density within and between lamellae, visual proteomics studies that use cryo-EM to establish changes in molecular organization within cells will require orders of magnitude more data than used in this study. One milestone would be to image enough data to represent one cellular volume, which for a small eukaryotic cells requires imaging approximately 100 lamellae. While data collection throughput on the TEM is fundamentally limited by the exposure time, this amount of data could be collected within 12 hours by improving the data acquisition scheme to perform all necessary calculations in parallel with actual exposure of the camera. Sample preparation using a FIB/SEM is also currently a bottleneck, but preparation of large lamellae with multiple cellular cross-sections using methods like WAFFLE {[}\protect\hyperlink{ref-17H2LWiIu}{28}{]} might allow sufficient throughput. As stated in the results, at least for 2DTM computation will remain challenging and approximately 17,000 GPU hours would be required for a 100 lamellae dataset.

@@

ER-HoxB8 cells were maintained in RPMI medium supplemented with 10\% FBS,
penicillin/streptomycin, SCF, and estrogen {[}\protect\hyperlink{ref-1B9Vt9eYu}{10}{]} at
37 °C and 5\% CO2. 120 h prior to grid preparation, cells were washed twice in PBS
and cultured in the same medium except without estrogen. Differentiation was
verified by staining with Hoechst-dye and inspection of nuclear morphology.
Cells were then counted and diluted to \(1\cdot10^6\) cells/ml. Grids (either 200
mesh copper grids, with a sillicone-oxide and 2 µm holes with a 2 µm spacing or
200 mesh gold grids with a thin gold film and 2 µm holes in 2 µm spacing) were
glow-discharged from both sides using a 15 mA for 45 s. 3.5 µl of cell suspension
was added to grids on the thin-film side and grids were blotted from the back
side using a GP2 cryoplunger (Leica, Wetzlar, Germany) for 8 s and rapidly plunged into liquid
ethane at -185 °C.

@@

Grids were loaded into an Aquilos 2 FIB/SEM (Thermo Fisher, Waltham, MA) instrument with a
stage cooled to -190 °C. Grids were sputter-coated with platinum for 15 s at 45 mA
and then coated with a layer of platinum-precursor by opening the GIS-valve for
45 s. An overview of the grid was created by montaging SEM images and isolated
cells at the center of gridsquares were selected for FIB-milling. Lamellae were
generated automatically using the AutoTEM software (Thermo Fisher), with the
following parameters:

@@

After data collection, a 20 s exposure at 2250x magnification of the lamella at
200 μm defocus was taken for visualization purposes. A Python script implementing
this procedure is available at \url{https://github.com/jojoelfe/deco_lace_template_matching_manuscript}.

@@

\begin{figure}
\hypertarget{fig:deco_lace_workflow}{%
\centering
\includegraphics{figures/deco_lace_workflow.png}
\caption{Workflow of DeCo-LACE processing}\label{fig:deco_lace_workflow}
}
\end{figure}
